---
title: "Education for Employment"
output:
  html_document: 
    theme: journal
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    toc_collapsed: TRUE
    css: ["style.css","MyFontsWebfontsKit.css"]
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
#install.packages("leaflet")
#install.packages("htmlwidgets")
#install.packages("spatstat.data")
#install.packages("spatstat")
library(leaflet)
library(htmlwidgets)
library(sf)
library(tidyverse)
library(tmap)
#library(spatstat)
library(maptools)
library(raster)
library(tidycensus)
library(sf)
library(jsonlite)
library(geojsonio)
library(hrbrthemes)
library(formattable)
library(scales)
library(ggrepel)
library(ggthemes)
library(kableExtra)
library(googlesheets4)


percent <- function(x, digits = 1, format = "f", ...) {
  paste0(formatC(100 * x, format = format, digits = digits, ...), "%")
}


```


### Overview

Education For Employment (EFE) is the leading nonprofit that trains youth and links them to jobs across the Middle East and North Africa (MENA).  This pivotal region is the hardest place on the planet for youth to get their first job – they are three times more likely to be unemployed than older adults. 

EFE is interested in the effectiveness of their programs, particularly whether graduates find stable employment. We have data on about 7,000 participants in almost 500 program cohorts spread across 8 countries. Participants bring diverse skills, interests, and backgrounds. Programs employ a variety of training models and placement policies. How well are different programs working, and for whom?


### Data

EFE has an Salesforce database that houses all information about the organization's programs, participants, and job placement and retention outcomes. The datasets used in this project include: 

* Participant demographics (program applications)
* Pre- and post- program participation surveys (job skills, confidence,…)
* Employment information at 3-month intervals for up to 2 years


### Data Pre-Processing

Initially, the data exports from Salesforce were pre-processed in the following ways:

* To protect PII, Contact ID, Program Name, Company Name, and Class Name were all anonymized by replacing the actual IDs/Names with new, numerical IDs. Lookup tables were sent to EFE so that they can bring the actual values back in if comapny, program, or class name are significant in the analysis.
* The applications dataset sometimes contained more than one application per person. Information from previous applications was deemed to be not as important, so only the most recent application was retained, and a new column containing the number of applications each person submitted was added. The application data was then joined to the contact data to create a more complete profile for each program participant.
* Any individuals who were not present in the Employment Status Check dataset were removed from the contact and pre/post survey datasets.



### Feature Engineering


```{r}

##Load Data 


##Must be authenticated with Google Sheets for this to work. Run line below and enter 0 in console to authenticate account
#gs4_auth()

Contacts<- read_sheet("https://docs.google.com/spreadsheets/d/1udA2SIU3UEptncprmma8xHe5RZWNm3bpxZUEb56CM5s/edit#gid=1903881109")
ESC<- read_sheet("https://docs.google.com/spreadsheets/d/1wtymKHb5Fk9BQ6un53HozXUKLIgyTeznHER039z1I5U/edit#gid=1677850102")
PrePost<- read_sheet("https://docs.google.com/spreadsheets/d/1e8P1YAPRN8xSKMSKA03SSWmdwdckSBErWnXiGIYID6Y/edit#gid=1255946150")







```

